\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{enumitem}


\title{Homework 3}
\author{Erik Gonzalez}

\begin{document}
\maketitle



\section*{Knights and Knaves:}



\begin{center}
\begin{tabular}{ |c|c|c|c| } 
 \hline
Case & P & Q \\ 
1 & Knight & Knight \\ 
2 & Knave & Knight \\ 
3 & Knight & Knave \\
4 & Knave & Knave \\
 \hline
\end{tabular}
\end{center}


\begin{itemize}
\item Knave: always tells lie
\item Knight: always tells truth 
\item If P is a Knight then whatever he/she says is true. If P is a Knave, whatever he/she says is false. If P was a Knight, case 1 can't happen, but 3 can. if P was a knave, both case 2 and 4 couldn't have happen because no matter what the knave says, will be false. Row 3 is the only possible case.
\end{itemize}


 
 
 
\begin{center}
\begin{tabular}{ |c|c|c|c| } 
 \hline
Case & P & Q \\ 
1 & Knight & Knight \\ 
2 & Knave & Knight \\ 
3 & Knight & Knave \\
4 & Knave & Knave \\
 \hline
\end{tabular}
\end{center}

 \begin{itemize}
\item  If P is a Knight then what he says would be true.. If P is a knave then what he says would be False. If P was a knight, case 1 and 3 couldn't have happened. If P was a knave, 4 couldn't have happened because whatever they say is a false. This leave case 2 being the only possible case.
\end{itemize}


\section*{Logical Identities: }

\begin{itemize}
\item $-(p \to (q \to p))$

 \begin{itemize}
\item Implication Law: $p \to q \equiv \neg q \vee p $. so, $ \neg (p \to (\neg q \vee p) $
\item Implication Law: $ \neg ( \neg p \vee (\neg q \vee p))$
\item De Morgans's Law: $\neg (a \vee b) \equiv \neg a \wedge \neg b,$  $\neg p \equiv  a $ $ and $ $(\neg q \vee p) \equiv b $ 
\begin{itemize}
\item $ \neg( \neg p) \wedge \neg ( \neg q \vee p)$
\end{itemize}
\item De Morgans's Law: $\neg (a \vee b)$ $ \equiv $ $\neg a \wedge \neg b,$
\begin{itemize}
\item $\neg q $ is a 
\item p is b 
\item $ \neg (\neg p) \wedge \neg(\neg q) \wedge \neg p$
\end{itemize}
\item double negation law: $\neg (\neg p ) \equiv p $ 
\begin{itemize}
\item $p \wedge q \wedge \neg p $
\item $p \wedge \neg p \wedge q $
\end{itemize}
\item $p \wedge \neg p \equiv F $, thus 
\item $FALSE \wedge q  \equiv q $
\end{itemize}
\end{itemize}



 \begin{itemize}
\item $-p((p \wedge  q) \to (q \vee p))$

 \begin{itemize}
 \item Let $(p \wedge q)$ be a, $(q \vee p)$ be b 
\item Implication Law, $ p \to q \equiv \neg p \vee q $
 \begin{itemize}
 \item $ \neg a \vee b $
 \end{itemize}
\item Substituting the values of a and b,
 \begin{itemize}
 \item $ \neg (p \wedge q) \vee ( q \vee p)$ 
 \end{itemize}
\item De Morgan's Law: $ \neg(p \wedge q) \equiv \neg p \vee  \neg q$ 
 \begin{itemize}
 \item $\neg p \vee \neg q \vee (q \vee p)$ ----------- 1
\end{itemize}
\item associative law: $(p \vee q) \vee r \equiv p \vee (q \vee r) $
\item Rearranging 1, 
 \begin{itemize}
 \item $(\neg p \vee p) \vee (\neg q \vee q)$
\end{itemize}
\item negation laws, $ p \vee \neg p \equiv T$
\item $TRUE \vee TRUE$ $ \equiv TRUE$
\end{itemize}
\end{itemize}


\section*{Logical Equivalences}


\begin{itemize}
\item $ p \to (q \to r)$ and $(p \wedge q) \to r $

 \begin{itemize}
\item Answer: $ p \to (q \to r)$ $ \ne  $ $(p \wedge q) \to r $

\vskip 0.1in

\vskip 0.1in



\end{itemize}

 \end{itemize} 







\begin{itemize}
\item $ p \to (q \to r)$ and $ (p \to q) \to r$

 \begin{itemize}
\item Answer: $ p \to (q \to r)$ $ \ne  $ $ (p \to q) \to r$




\end{itemize}

 \end{itemize} 







\section*{Logical Consequence}



\begin{itemize}
\item 1. 

Jimmy is smart
\vskip 0.1in
Smart People are rich
\vskip 0.1in
 \textbf{----------------------------}
\vskip 0.1in
Jimmy is Rich 
 \vskip 0.1in
 
 
 \textbf{Answer:} Valid: Jimmy is rich 
 
 Jimmy is smart
\vskip 0.1in
Jimmy then smart
\vskip 0.1in
Jimmy $\to$ smart

Smart People are rich
\vskip 0.1in
Smart then rich 
\vskip 0.1in
Smart $\to$ rich 
\vskip 0.1in
Jimmy $\to$ rich $\equiv$ Jimmy is rich 

  


 \end{itemize} 
    
    
    
    
   
 
\begin{itemize}
\item 2. 

Islands are surrounded by water
\vskip 0.1in 
Puerto Rico is surrounded by water
\vskip 0.1in 
\textbf{--------------------------------------}
\vskip 0.1in 
Puerto Rico is an island
\vskip 0.1in


 \textbf{Answer:}  Not Valid: We cannot say Puerto Rico is an island is always true. 
 
 Islands are surrounded by water
 \vskip 0.1in
 Islands $\to$ surrounded by water
 \vskip 0.1in
  Puerto Rico is surrounded by water
 \vskip 0.1in
 Puerto Rico $\to$ surrounded by water
  \vskip 0.1in
  Puerto Rico is not always an island 
  
   
   
 \end{itemize} 






\end{document}